\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-03-17 jasa/paper-jasa.tex
remove redundant equations

Finally, we come to the log-likelihood curve
%shown in Figure~\ref{fig:likelihood}
\begin{equation}
\log \Lambda( \gamma ) =  \log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} \;
\end{equation}
for the dataset $\mathcal{D}$ shown in  Fig.~\ref{fig:1a}.
By exploiting the fact that
\begin{equation}
\log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} = \log \frac{ p({\cal D} | \gamma ) }{ p({\cal D} | \gamma = 0 ) } - \log \frac{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma) }{  p({\cal D} | \gamma = 0 ) },
\end{equation}
the generalized likelihood ratio can be computed by evaluating
both terms with respect to a common reference $\gamma=0$
%reusing the decomposed neural network-based approximation from Figure~\ref{fig:1c}.
%In particular, the maximum likelihood estimator (MLE) in the second term
%can be found
as outlined in Section~\ref{sec:generalized-likelihood-ratio}.
Figure~\ref{fig:likelihood} shows that the exact likelihood curve is very well
approximated by the method, confirming that
even when the raw classifier does a poor job at modeling the $s^*(\mathbf{x})$, a good approximations of the
likelihood ratio can still be obtained by calibrating $s(\mathbf{x})$ (and by decomposing
the mixture, if possible).

@@

    \begin{subfigure}[t]{0.4\textwidth}
        \includegraphics[clip, trim=0.5cm 0.5cm 0.5cm 0.5cm,width=\textwidth]{figures/fig2a.pdf}
        \caption{Exact vs. approximated MLEs.}
        \label{fig:2a}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.4\textwidth}
        \includegraphics[clip, trim=0.5cm 0.5cm 0.5cm 0.5cm,width=\textwidth]{figures/fig2b.pdf}
        \caption{$p(-2 \log \Lambda(\gamma=0.05) \, | \, \gamma=0.05)$}
        \label{fig:2b}
    \end{subfigure}

@@

The observations ${\cal D}$ represented in Fig.~\ref{fig:4}
are random samples with $\alpha=1$ and $\beta=-1$.
Our goal is to infer the values $\alpha$ and $\beta$ based on ${\cal D}$.
We construct the log-likelihood ratio
\begin{equation}\label{eqn:obj-ex2}
    -2 \log \Lambda(\alpha, \beta ) = -2 \log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }
\end{equation}
that we calculate by exploiting the fact that
\begin{equation}\label{eqn:obj-ex2-expanded}
\log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) } = \log \frac{p({\cal D} | \alpha, \beta)}{ p({\cal D} | \alpha=0, \beta=0) } - \log \frac{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }{ p({\cal D} | \alpha=0, \beta=0)}.
\end{equation}
Following the procedure described in Section~\ref{sec:param-clf}, we
% to reduce
%computational requirements, both terms in the right hand side of Eqn.~\ref{eqn:obj-ex2-expanded} are evaluated using
build a single 2-layer neural network (with 5+2 inputs and one output node) to form the parameterized classifier $s(\mathbf{x}; \theta_0, \theta_1)$ and fix $\theta_1=(\alpha=0, \beta=0)$.
Since the generative model is not expensive, the classifier output is calibrated
on-the-fly with histograms for  every candidate parameter pair $(\alpha,\beta)$.

>>>>

Finally, we come to the log-likelihood curve
%shown in Figure~\ref{fig:likelihood}
\begin{equation}
\log \Lambda( \gamma ) =  \log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} \;
\end{equation}
for the dataset $\mathcal{D}$ shown in  Fig.~\ref{fig:1a}.
By exploiting Eq.~\ref{eq:theta_1_independence},
% fact that
%\begin{equation}
%\log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} = \log \frac{ p({\cal D} | \gamma ) }{ p({\cal D} | \gamma = 0 ) } - \log \frac{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma) }{  p({\cal D} | \gamma = 0 ) },
%\end{equation}
the generalized likelihood ratio can be computed by evaluating
both terms with respect to a common reference $\gamma=0$
%reusing the decomposed neural network-based approximation from Figure~\ref{fig:1c}.
%In particular, the maximum likelihood estimator (MLE) in the second term
%can be found
as outlined in Section~\ref{sec:generalized-likelihood-ratio}.
Figure~\ref{fig:likelihood} shows that the exact likelihood curve is very well
approximated by the method, confirming that
even when the raw classifier does a poor job at modeling the $s^*(\mathbf{x})$, a good approximations of the
likelihood ratio can still be obtained by calibrating $s(\mathbf{x})$ (and by decomposing
the mixture, if possible).

@@

    \begin{subfigure}[t]{0.475\textwidth}
        \includegraphics[clip, trim=0.5cm 0.5cm 0.5cm 0.5cm,width=\textwidth]{figures/fig2a.pdf}
        \caption{Exact vs. approximated MLEs.}
        \label{fig:2a}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.475\textwidth}
        \includegraphics[clip, trim=0.5cm 0.5cm 0.5cm 0.5cm,width=\textwidth]{figures/fig2b.pdf}
        \caption{$p(-2 \log \Lambda(\gamma=0.05) \, | \, \gamma=0.05)$}
        \label{fig:2b}
    \end{subfigure}

@@

The observations ${\cal D}$ represented in Fig.~\ref{fig:4}
are random samples with $\alpha=1$ and $\beta=-1$.
Our goal is to infer the values $\alpha$ and $\beta$ based on ${\cal D}$.
We construct the log-likelihood ratio
\begin{equation}\label{eqn:obj-ex2}
    -2 \log \Lambda(\alpha, \beta ) = -2 \log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }
\end{equation}
that we calculate via Eqn.~\ref{eq:theta_1_independence}.
%by exploiting the fact that
%\begin{equation}\label{eqn:obj-ex2-expanded}
%\log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) } = \log \frac{p({\cal D} | \alpha, \beta)}{ p({\cal D} | \alpha=0, \beta=0) } - \log \frac{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }{ p({\cal D} | \alpha=0, \beta=0)}.
%\end{equation}
Following the procedure described in Section~\ref{sec:param-clf}, we
% to reduce
%computational requirements, both terms in the right hand side of Eqn.~\ref{eqn:obj-ex2-expanded} are evaluated using
build a single 2-layer neural network (with 5+2 inputs and one output node) to form the parameterized classifier $s(\mathbf{x}; \theta_0, \theta_1)$ and fix $\theta_1=(\alpha=0, \beta=0)$.
Since the generative model is not expensive, the classifier output is calibrated
on-the-fly with histograms for  every candidate parameter pair $(\alpha,\beta)$.
